---
title: "frequency model brms"
output: html_document
date: "2023-04-17"
---

```{r}
library(pacman)
p_load(data.table, janitor, accelerateR, brms, tidyverse, ggplot2)
```

# real data
## Acerodon jubatus
### load data
```{r}
load("../../../../Dropbox/MPI/Wingbeat/data/Frequency_Ajubatus_2010.robj")
```

###Summarise data
```{r}
Freq$day_bat_species <- paste(date(Freq$sunset), Freq$bat, Freq$species)
day_bat_species <- table(Freq$day_bat_species)
length(day_bat_species)
hist(day_bat_species)
```

### filter by frequency and signal amplitude
```{r}
plot(Freq$frequency, Freq$rms)

#plot(Freq$frequency, Freq$amp, col = rgb(0,0,0,.1))
freq_threshold_low <- 1.5
freq_threshold_high <- 5
rms_threshold <- 100
abline(v = freq_threshold_low)
abline(v = freq_threshold_high)
abline(h = rms_threshold)

# hist(Freq$frequency, breaks = 100)
# hist(Freq$amp, breaks = 100)
```

#### frequency and amplitude threshold
```{r}
with(Freq[Freq$frequency > freq_threshold_low & Freq$frequency < freq_threshold_high & 
            Freq$rms > rms_threshold,], 
     plot(frequency, amp, col = factor(behavior), cex = 0.5))

clean_freq <- Freq[Freq$frequency > freq_threshold_low & Freq$frequency < freq_threshold_high & 
            Freq$rms > rms_threshold,]

with(clean_freq, plot(frequency, rms_filter, col = behavior %>% factor))

```


#### remove points before sunset
```{r}
hist(clean_freq$hours_since_sunset/24)
clean_freq <- clean_freq[clean_freq$hours_since_sunset > 0,]

```

#### clean up nights with little acivity
```{r}
clean_freq$date <- date(clean_freq$sunset)
clean_freq$bat %>% table %>% length

day_bat_species <- clean_freq$day_bat_species %>% table
clean_freq$day_bat_species %>% table %>% length
clean_freq$day_bat_species %>% table %>% hist(breaks = 20)

keep <- day_bat_species[which(day_bat_species %>% as.numeric > 20)] %>% names()
clean_freq <- clean_freq[which(clean_freq$day_bat_species %in% keep),]

clean_freq$bat %>% table %>% length
clean_freq$day_bat_species %>% table %>% length()

ggplot(clean_freq, aes(y = frequency, x = hours_since_sunset, group = day_bat_species))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~day_bat_species)
```


### run model
```{r}
fit_aj <- brm(frequency ~ hours_since_sunset + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq %>% na.omit(), 
            cores = 20)
summary(fit_aj)
plot(fit_aj, ask = FALSE)

save(fit_aj, Freq, clean_freq, freq_threshold_low, freq_threshold_high, amp_threshold,
     file = "../../../../Dropbox/MPI/Wingbeat/data/Frequency_Ajubatus_model.robj")

# load("../../../../Dropbox/MPI/Wingbeat/data/Frequency_Phastatus_model.robj")
```



# real data
## eobs
### load data
```{r}
load("../../../../Dropbox/MPI/Wingbeat/data/Frequency_eobs.robj")
```

# clean
```{r}
summary(Frequency)

Frequency <- Frequency[Frequency$hours_since_sunset > 0 & 
                        Frequency$frequency > 1 &
                         Frequency$amp > 50 &
                         Frequency$behavior != "resting",] %>% na.omit()
```

# make sure there are enough observations per bat night
```{r}
Frequency$day_bat_species <- paste(date(Frequency$sunset), Frequency$bat, Frequency$species)
day_bat_species <- table(Frequency$day_bat_species)

keep <- day_bat_species[which(day_bat_species %>% as.numeric > 30)] %>% names()

clean_freq <- Frequency[which(Frequency$day_bat_species %in% keep),]
clean_freq$day_bat_species %>% table
```


```{r}
clean_freq$date <- date(clean_freq$sunset)
clean_freq %>% head
summary(clean_freq)

```
# summarise data
```{r}
clean_freq

clean_freq$bat[clean_freq$species=="Ajubatus" &
                                clean_freq$behavior == "commuting" &
                                clean_freq$frequency < 5] %>% table %>% length
clean_freq$day_bat_species[clean_freq$species=="Ajubatus" &
                                clean_freq$behavior == "commuting" &
                                clean_freq$frequency < 5] %>% table %>% length

clean_freq$bat[clean_freq$species=="Ehelvum"] %>% table %>% length
clean_freq$day_bat_species[clean_freq$species=="Ehelvum"] %>% table %>% length

clean_freq$bat[clean_freq$species=="Plylei"] %>% table %>% length
clean_freq$day_bat_species[clean_freq$species=="Plylei"] %>% table %>% length

```


# plot data
```{r}
ggplot(clean_freq[clean_freq$hours_since_sunset > 0,], 
       aes(x = round(hours_since_sunset,0), y = frequency, col = bat, group = bat))+
  geom_smooth(method = "lm")+ xlab("Hours since sunset")+ ylab("Dominant Wingbeat Frequency (Hz)")+
  facet_wrap(~species)+theme_bw()+theme(legend.position = "none")
# 
# ggplot(clean_freq, 
#        aes(x = frequency, y = amp, col = bat, group = bat))+
#   geom_point(alpha = 0.1)+
#   # geom_smooth(method = "lm")+
#   facet_wrap(~species)+theme_bw()+theme(legend.position = "none")
# 
# ggplot(clean_freq[clean_freq$hours_since_sunset > 0 &
#                  clean_freq$behavior == "commuting",], 
#        aes(x = frequency, y = amp, col = bat, group = bat))+
#   geom_smooth(method = "lm")+
#   facet_wrap(~species)+theme_bw()+theme(legend.position = "none")
```


# run models
## Acerodon
```{r}
clean_freq$day_bat_species[clean_freq$species=="Ajubatus"&
                                #clean_freq$behavior == "commuting" &
                             clean_freq$frequency > 1 &   
                             clean_freq$frequency < 5] %>% table
ggplot(clean_freq[clean_freq$species=="Ajubatus"&
                                #clean_freq$behavior == "commuting" &
                    clean_freq$frequency > 1 &            
                    clean_freq$frequency < 5,] %>% na.omit(),
       aes(hours_since_sunset, frequency, col = bat))+ geom_point()+
  geom_smooth(method = "lm")+facet_wrap(~bat+date, scales = "free")

fit_aj <- brm(frequency ~ round(hours_since_sunset, 0) + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Ajubatus" &
                                #clean_freq$behavior == "commuting" &
                                clean_freq$frequency > 1 &
                                clean_freq$frequency < 5,] %>% na.omit(), 
            cores = 20)
summary(fit_aj)
plot(fit_aj, ask = FALSE)
```

### amplitude frequency
```{r}
ggplot(clean_freq[clean_freq$species=="Ajubatus"&
                                clean_freq$behavior == "commuting",],
                                #clean_freq$frequency < 5
       aes(hours_since_sunset, amp, col = bat))+ geom_point()+
  geom_smooth(method = "lm")#+facet_wrap(~bat+date, scales = "free")
```


## Eidolon
```{r}
ggplot(clean_freq[clean_freq$species=="Ehelvum"&
                                clean_freq$behavior == "commuting" &
                    clean_freq$frequency > 3 &            
                    clean_freq$frequency < 6,],
       aes(hours_since_sunset, frequency, col = bat))+ geom_point()+
  geom_smooth(method = "lm")+facet_wrap(~bat+date, scales = "free")

fit_eh <- brm(frequency ~ round(hours_since_sunset, 0) + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Ehelvum",] %>% na.omit(), 
            cores = 20)
summary(fit_eh)
plot(fit_eh, ask = FALSE)
```

## Pteropus
```{r}
fit_pl <- brm(frequency ~ round(hours_since_sunset, 0) + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Plylei" &
                                clean_freq$behavior == 'commuting',] %>% na.omit(), 
            cores = 20)
summary(fit_pl)
plot(fit_pl, ask = FALSE)
```


```{r}
# fit_eobs <- brm(frequency ~ round(hours_since_sunset, 0) + (1+date|bat+species),
#             family = gaussian,
#             # prior = c(prior(normal(2.5, sd = 0.5)))
#             data = clean_freq %>% na.omit(), 
#             cores = 20)
# summary(fit_eobs)
# plot(fit_eobs, ask = FALSE)
```

```{r}
save(fit_aj, fit_eh, fit_pl, Frequency, clean_freq, 
     file = "../../../../Dropbox/MPI/Wingbeat/data/Frequency_eobs_commuting_hours_models.robj")
```

