---
title: "frequency model brms"
output: html_document
date: "2023-04-17"
---

Can we model changes in wingbeat frequency with time since sunset?

# load libraries
```{r}
library(pacman)
p_load(data.table, janitor, accelerateR, brms, tidyverse, ggplot2)
```

# real data
## Phyllostomus hastatus
### load data
```{r}
load("../../../../Dropbox/MPI/Wingbeat/data/Frequency_Phastatus_2022.robj")
```

###Summarise data
```{r}
Freq$day_bat_species <- paste(date(Freq$sunset), Freq$bat, Freq$species)
day_bat_species <- table(Freq$day_bat_species)
length(day_bat_species)
hist(day_bat_species)
```

### filter by frequency and signal amplitude
```{r}
plot(Freq$frequency, Freq$amp, col = rgb(0,0,0,.1))
freq_threshold <- 5
amp_threshold <- 0.075
abline(v = freq_threshold)
abline(h = amp_threshold)

hist(Freq$frequency, breaks = 100)
hist(Freq$amp, breaks = 100)
```

#### frequency and amplitude threshold
```{r}
with(Freq[Freq$frequency > freq_threshold & Freq$amp > amp_threshold,], 
     plot(frequency, amp, col = factor(behavior), cex = 0.1))

clean_freq <- Freq[Freq$frequency > freq_threshold & Freq$amp > amp_threshold,]

```


#### remove points before sunset
```{r}
hist(clean_freq$hours_since_sunset/24)
clean_freq <- clean_freq[clean_freq$hours_since_sunset > 0,]

```

#### clean up nights with little acivity
```{r}
clean_freq$date <- date(clean_freq$sunset)
clean_freq$bat %>% table %>% length

clean_freq$day_bat_species %>% table %>% length
clean_freq$day_bat_species %>% table %>% hist(breaks = 20)

keep <- day_bat_species[which(day_bat_species %>% as.numeric > 50)] %>% names()
clean_freq <- clean_freq[which(clean_freq$day_bat_species %in% keep),]

clean_freq$bat %>% table %>% length
clean_freq$day_bat_species %>% table %>% length()

ggplot(clean_freq, aes(y = frequency, x = hours_since_sunset, group = day_bat_species))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~day_bat_species)
```


### run model
```{r}
fit_ph <- brm(frequency ~ hours_since_sunset + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Nlasiopterus",] %>% na.omit(), 
            cores = 20)
summary(fit_ph)
plot(fit_ph, ask = FALSE)

save(fit_ph, Freq, clean_freq, freq_threshold, amp_threshold,
     file = "../../../../Dropbox/MPI/Wingbeat/data/Frequency_Phastatus_model.robj")
```


## Myotis vivesi
```{r}
fit_mv <- brm(frequency ~ round(hours_since_sunset, 0) + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Mvivesi" &
                                clean_freq$behavior == "commuting" &
                                clean_freq$frequency > 5.5 &
                                clean_freq$frequency < 7.5,] %>% na.omit(), 
            cores = 20)
summary(fit_mv)
plot(fit_mv, ask = FALSE)
save(fit_mv, Frequency, clean_freq, 
     file = "../../../../Dropbox/MPI/Wingbeat/data/Frequency_vesper_commuting_hours_models.robj")
```

## Nyctalus
### load data
```{r}
load("../../../../Dropbox/MPI/Wingbeat/data/Frequency_Nlasiopterus_2018.robj")
```

###Summarise data
```{r}
Freq$day_bat_species <- paste(date(Freq$sunset), Freq$bat, Freq$species)
day_bat_species <- table(Freq$day_bat_species)
length(day_bat_species)
hist(day_bat_species)
```

### filter by frequency and signal amplitude
```{r}
plot(Freq$frequency, Freq$amp, col = rgb(0,0,0,.1))

hist(Freq$frequency, breaks = 100)
hist(Freq$amp, breaks = 100)
```


```{r}
# library(dbscan)
# dbscan_res <- dbscan(Freq[,c("frequency", "amp")], eps = 0.2, minPts = 50)
# plot(Freq[,c("frequency", "amp")], col = dbscan_res$cluster + 1)
```


#### frequency and amplitude threshold
```{r}
freq_threshold <- 5
amp_threshold <- 0.5

with(Freq[Freq$frequency > freq_threshold & Freq$amp > amp_threshold,], 
     plot(frequency, amp, col = factor(behavior), cex = 0.1))

clean_freq <- Freq[Freq$frequency > freq_threshold & Freq$amp > amp_threshold,]

```

#### remove points before sunset
```{r}
clean_freq <- clean_freq[clean_freq$hours_since_sunset > 0,]

```


#### clean up nights with little acivity
```{r}
clean_freq$date <- date(clean_freq$sunset)
clean_freq$bat %>% table %>% length

clean_freq$day_bat_species %>% table %>% length
clean_freq$day_bat_species %>% table %>% hist(breaks = 20)

keep <- day_bat_species[which(day_bat_species %>% as.numeric > 50)] %>% names()
clean_freq <- clean_freq[which(clean_freq$day_bat_species %in% keep),]

clean_freq$bat %>% table %>% length
clean_freq$day_bat_species %>% table %>% length()

ggplot(clean_freq, aes(y = frequency, x = hours_since_sunset, group = day_bat_species))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~day_bat_species)
```



### run model
```{r}
fit_nl <- brm(frequency ~ hours_since_sunset + (1+date|bat),
            family = gaussian,
            # prior = c(prior(normal(2.5, sd = 0.5)))
            data = clean_freq[clean_freq$species=="Nlasiopterus",] %>% na.omit(), 
            cores = 20)
summary(fit_nl)
plot(fit_nl, ask = FALSE)

save(fit_nl, Freq, clean_freq, freq_threshold, amp_threshold,
     file = "../../../../Dropbox/MPI/Wingbeat/data/Frequency_Nlasiopterus_model.robj")
```

```{r}

```

